---
title: "p8105_hw2_st3117"
author: "Sha Tao"
date: "September 27, 2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1_1 NYC Transit Table Cleaning

```{r Problem 1_1}
# import and clean data
transit = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  gather(key = routes_number, value = routes_name, route1:route11) %>% 
  separate(routes_number, into = c("remove1", "routes_number"), sep = 5) %>% 
  filter(!is.na(routes_name)) %>% 
  mutate(entry = ifelse(entry == "YES", yes = TRUE, no = FALSE)) %>% 
  select(line:station_longitude, routes_number, routes_name, entry, vending, entrance_type, ada) %>% 
  arrange(line, station_name)

transit

```

The dataset "transit" is a `r ncol(transit)` by `r nrow(transit)` table.\
It contains variables including line, station_name, latitude, longitude, routes_number, routes_name, entry, vending, entrance_type and ADA compliance.\
I started by importing the dataset and clean the variable names. Then, I used 'gather' function doing a wide to long transpose; used 'separate' function removing characters "route" from routes_number; used 'filter' function removing “NA”s in routes_name; used 'ifelse' function converting entry to a logical variable; seleted the 10 variables as requested; and arrange them by line and station_name.\

This is a tidy dataset.

## Problem 1_2 Answering questions

* How many distinct stations are there?\
There are `r count(distinct(transit, line, station_name))` distinct stations.\

* How many stations are ADA compliant?\
`r count(filter(transit, ada == TRUE) %>% distinct(line, station_name))` stations are ADA compliant.\

* What proportion of station entrances / exits without vending allow entrance?\
`r round(count(filter(transit, vending == "NO" & entry == TRUE) %>% distinct(line, station_name))*100 /  count(filter(transit, vending == "NO") %>% distinct(line, station_name)), 2)`% station entrances / exits without vending allow entrance.

## Problem 1_3
```{r}
filter(transit, routes_name == "A") %>% 
  distinct(line, station_name, routes_number, routes_name) %>% 
  nrow()

count(filter(transit, routes_name == "A") %>% 
  distinct(line, station_name, routes_number, routes_name))

filter(transit, ada == TRUE & routes_name == "A") %>% 
  distinct(line, station_name, routes_number, routes_name) %>% 
  nrow()

count(filter(transit, ada == TRUE & routes_name == "A") %>% 
  distinct(line, station_name, routes_number, routes_name))
```


## Problem 2_1
```{r}

trash = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                     sheet = "Mr. Trash Wheel", range = readxl::cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  filter(!is.na(date)) %>% 
  mutate(
    sports_balls = round(sports_balls) %>% 
                   as.integer(sports_balls)
    )

#    year = ifelse(!is.na(year), year, substring(date, 1, 4)),

trash

    

```


## Problem 2_2
```{r}

prec_2016 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2016 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2016")
                              
prec_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2017") %>% 
  filter(!is.na(total))

precipitation = 
  bind_rows(prec_2016, prec_2017) %>%
  select(year, month, total) %>% 
  mutate(month = month.name[month])

precipitation
  

```


```{r}
#filter(trash, year == 2016) %>% 
#  pull(sports_balls) %>% 
#  median()

#median(filter(trash, year == 2016) %>% trash$sports_balls)

count(trash)
count(precipitation)
sum(prec_2017$total)

```

`r median(filter(trash, year == 2016) %>% pull(sports_balls))`

`r nrow(trash)`

`r nrow(precipitation)`

`r sum(prec_2017$total)`



```{r}

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("remove2", "county"), sep = 5) %>% 
  select(year, state = locationabbr, county, response, data_value) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  select(year:county, excellent, very_good, good, fair, poor) %>% 
  mutate(above_good = excellent + very_good)
  
brfss
```


```{r}
#How many unique locations are included in the dataset?
location_distinct = distinct(brfss, state, county)

state_distinct = distinct(brfss, state)

count(location_distinct, state) %>% 
arrange(-n) %>% 
head(1)

count(brfss, state) %>% 
arrange(-n) %>% 
head(1)

brfss_2002 = filter(brfss, year == 2002)

pull(brfss_2002, excellent) %>% 
  median(na.rm = TRUE)

filter(brfss, year == 2002) %>% 
  pull(excellent) %>% 
  median(na.rm = TRUE)

pull(brfss, excellent) %>% 
  median(na.rm = TRUE)


ggplot(brfss_2002, aes(x = excellent)) + 
  geom_histogram()

filter(brfss, year == 2002) %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram()


brfss_ny = filter(brfss, county == "New York County" | county == "Queens County") %>% 
  arrange(county, year)

ggplot(brfss_ny, aes(x = year, y = excellent)) + 
  geom_point(aes(color = county))

```



