---
title: "p8105_hw2_st3117"
author: "Sha Tao"
date: "September 27, 2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1_1 NYC Transit Data Cleaning

```{r Problem 1_1, message = FALSE}

# import and clean data
transit = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  gather(key = routes_number, value = routes_name, route1:route11) %>% 
  separate(routes_number, into = c("remove1", "routes_number"), sep = 5) %>% 
  filter(!is.na(routes_name)) %>% 
  mutate(entry = ifelse(entry == "YES", yes = TRUE, no = FALSE)) %>% 
  select(line:station_longitude, routes_number, routes_name, entry, vending, entrance_type, ada) %>% 
  arrange(line, station_name)

```

The dataset "transit" is a `r ncol(transit)` by `r nrow(transit)` table.\
It contains variables including line, station_name, latitude, longitude, routes_number, routes_name, entry, vending, entrance_type and ADA compliance.\
I started by importing the dataset and clean the variable names. Then, I used 'gather' function doing a wide to long transpose; used 'separate' function removing characters "route" from routes_number; used 'filter' function removing “NA”s in routes_name; used 'ifelse' function converting entry to a logical variable; seleted the 10 variables as requested; and arrange them by line and station_name.\
</br>
This is a tidy dataset.

## Problem 1_2 Answer Questions

* How many distinct stations are there?\
There are `r count(distinct(transit, line, station_name))` distinct stations.\
</br>
* How many stations are ADA compliant?\
`r count(filter(transit, ada == TRUE) %>% distinct(line, station_name))` stations are ADA compliant.\
</br>
* What proportion of station entrances / exits without vending allow entrance?\
`r round(count(filter(transit, vending == "NO" & entry == TRUE) %>% distinct(line, station_name))*100 /  count(filter(transit, vending == "NO") %>% distinct(line, station_name)), 2)`% station entrances / exits without vending allow entrance.

## Problem 1_3 Reformat Data and Answer Questions

* How many distinct stations serve the A train?\
There are `r count(filter(transit, routes_name == "A") %>% distinct(line, station_name, routes_number, routes_name))` distinct stations serve the A train.\
</br>
* Of the stations that serve the A train, how many are ADA compliant?\
`r count(filter(transit, ada == TRUE & routes_name == "A") %>% distinct(line, station_name, routes_number, routes_name))` Of the stations that serve the A train are ADA compliant.

## Problem 2_1 Mr. Trash Wheel Data Cleaning

```{r Problem 2_1}

# import and clean data
trash = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                     sheet = "Mr. Trash Wheel", range = readxl::cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(
    sports_balls = round(sports_balls) %>% 
                   as.integer(sports_balls)
    )

```


## Problem 2_2 Precipitation Data Cleaning

```{r Problem 2_2}

# import and clean data for 2016
prec_2016 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2016 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2016")

# import and clean data for 2017                             
prec_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2017") 
  
# note: the previous dataset contains NAs for precipitation thus need ' %>% filter(!is.na(total))', but this one is complete.

# combine the 2016 and 2017 datasets, and change month type
precipitation = 
  bind_rows(prec_2016, prec_2017) %>%
  select(year, month, total) %>% 
  mutate(month = month.name[month])

```


## Problem 2_3 Describe Two Datasets

```{r Problem 2_3}

# Find the highest total precipitation
arrange(precipitation, -total) %>% 
  head(1)

```

After cleaning, the dataset ‘trash’ is a `r ncol(trash)` by `r nrow(trash)` (observations) table. It includes the trash weight and trash types collected by the Mr. Trash Wheel from 2017 to 2018. The key variables are weight_tons and homes_powered. From the dataset, we know that each dumpster collected `r round(mean(trash$weight_tons, na.rm = TRUE), 2)` tons on average, and they powered `r round(sum(trash$homes_powered), 2)` homes in total for the two years.\
The dataset ‘precipitation’ is a `r ncol(precipitation)` by `r nrow(precipitation)` (observations) table. It records the monthly total precipitation from 2016 to 2017. The key variables are total_precipitation. From the dataset, we know that the monthly total percipitation is`r round(mean(precipitation$total, na.rm = TRUE), 2)` inches on average, and the highest total monthly precipitation occured in July, 2017 with 7.09 inches.\

## Problem 2_4 Answer Questions

```{r Problem 2_4}
filter(trash, year == 2016) %>% 
  pull(sports_balls) %>% 
  median(na.rm = TRUE)

```

*  what was the total precipitation in 2017?\
The total precipitation in 2017 are `r sum(prec_2017$total)` inches.\
</br>
* What was the median number of sports balls in a dumpster in 2016?\
The median number of sports balls in a dumpster in 2016 are `r median(filter(trash, year == 2016) %>% pull(sports_balls), na.rm = TRUE)`.



```{r}

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("remove2", "county"), sep = 5) %>% 
  select(year, state = locationabbr, county, response, data_value) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  select(year:county, excellent, very_good, good, fair, poor) %>% 
  mutate(above_good = excellent + very_good)
  
brfss
```


```{r}
#How many unique locations are included in the dataset?
location_distinct = distinct(brfss, state, county)

state_distinct = distinct(brfss, state)

count(location_distinct, state) %>% 
arrange(-n) %>% 
head(1)

count(brfss, state) %>% 
arrange(-n) %>% 
head(1)

brfss_2002 = filter(brfss, year == 2002)

pull(brfss_2002, excellent) %>% 
  median(na.rm = TRUE)

filter(brfss, year == 2002) %>% 
  pull(excellent) %>% 
  median(na.rm = TRUE)

pull(brfss, excellent) %>% 
  median(na.rm = TRUE)


ggplot(brfss_2002, aes(x = excellent)) + 
  geom_histogram()

filter(brfss, year == 2002) %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram()


brfss_ny = filter(brfss, county == "New York County" | county == "Queens County") %>% 
  arrange(county, year)

ggplot(brfss_ny, aes(x = year, y = excellent)) + 
  geom_point(aes(color = county))

```



