---
title: "p8105_hw2_st3117"
author: "Sha Tao"
date: "September 27, 2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1_1 NYC Transit Data Cleaning

```{r Problem 1_1, message = FALSE}

# import and clean data
transit = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  gather(key = routes_number, value = routes_name, route1:route11) %>% 
  separate(routes_number, into = c("remove1", "routes_number"), sep = 5) %>% 
  filter(!is.na(routes_name)) %>% 
  mutate(entry = ifelse(entry == "YES", yes = TRUE, no = FALSE)) %>% 
  select(line:station_longitude, routes_number, routes_name, entry, vending, entrance_type, ada) %>% 
  arrange(line, station_name)

```

The dataset "transit" is a `r ncol(transit)` by `r nrow(transit)` table.\
It contains variables including line, station_name, latitude, longitude, routes_number, routes_name, entry, vending, entrance_type and ADA compliance.\
I started by importing the dataset and clean the variable names. Then, I used 'gather' function doing a wide to long transpose; used 'separate' function removing characters "route" from routes_number; used 'filter' function removing “NA”s in routes_name; used 'ifelse' function converting entry to a logical variable; seleted the 10 variables as requested; and arranged them by line and station_name.\
</br>
This is a tidy dataset.

## Problem 1_2 Answer Questions about Transit

* How many distinct stations are there?\
There are `r count(distinct(transit, line, station_name))` distinct stations.\
</br>
* How many stations are ADA compliant?\
`r count(filter(transit, ada == TRUE) %>% distinct(line, station_name))` stations are ADA compliant.\
</br>
* What proportion of station entrances / exits without vending allow entrance?\
`r round(count(filter(transit, vending == "NO" & entry == TRUE) %>% distinct(line, station_name))*100 /  count(filter(transit, vending == "NO") %>% distinct(line, station_name)), 2)`% station entrances / exits without vending allow entrance.

## Problem 1_3 Reformat Data Transit and Answer Questions

* How many distinct stations serve the A train?\
There are `r count(filter(transit, routes_name == "A") %>% distinct(line, station_name, routes_number, routes_name))` distinct stations serve the A train.\
</br>
* Of the stations that serve the A train, how many are ADA compliant?\
`r count(filter(transit, ada == TRUE & routes_name == "A") %>% distinct(line, station_name, routes_number, routes_name))` stations that serve the A train are ADA compliant.

## Problem 2_1 Mr. Trash Wheel Data Cleaning

```{r Problem 2_1, warning = FALSE}

# import and clean data
trash = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                     sheet = "Mr. Trash Wheel", range = readxl::cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(
    sports_balls = round(sports_balls) %>% 
                   as.integer(sports_balls)
    )

# try the tidy form, but won't use this dataset for the analysis later
trash_tidy = trash %>% 
  gather(key = trash_type, value = amount, plastic_bottles:sports_balls) %>% 
  separate(trash_type, into = c("trash_type", "unit"), sep = "_") %>% 
  mutate(unit = ifelse(is.na(unit), yes = "units", no = unit)) %>% 
  arrange(dumpster, year, month) %>% 
  select(dumpster:volume_cubic_yards, trash_type:amount, homes_powered)
```

Note: I tried to make the dataset 'trash' tidy by transposing different trash_type to a long form. However, this trash_tidy dataset contains 1-to-many relations, we can either make relational tables or using 'group_by' for the data analysis, which is not convenient at all. Thus I'll stick to the dataset 'trash'.


## Problem 2_2 Precipitation Data Cleaning

```{r Problem 2_2}

# import and clean data for 2016
prec_2016 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2016 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2016")

# import and clean data for 2017                             
prec_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2017") 
  
# note: the previous dataset contains NAs for precipitation thus need ' %>% filter(!is.na(total))', but this one is complete.

# combine the 2016 and 2017 datasets, and change month type
precipitation = 
  bind_rows(prec_2016, prec_2017) %>%
  select(year, month, total) %>% 
  mutate(month = month.name[month])

```


## Problem 2_3 Describe Two Datasets Trash and Precipitation

```{r Problem 2_3}

# Find the highest total precipitation
arrange(precipitation, -total) %>% 
  head(1)

```

After cleaning, the dataset ‘trash’ is a `r ncol(trash)` by `r nrow(trash)` (observations) table. It includes the trash weight and trash types collected by the Mr. Trash Wheel from 2014 to 2018. The key variables are weight_tons and homes_powered. From the dataset, we know that each dumpster collected `r round(mean(trash$weight_tons, na.rm = TRUE), 2)` tons on average, and they powered `r round(sum(trash$homes_powered),-2)` homes in total for the past four years.\
The dataset ‘precipitation’ is a `r ncol(precipitation)` by `r nrow(precipitation)` (observations) table. It records the monthly total precipitation from 2016 to 2017. The key variables are total_precipitation. From the dataset, we know that the monthly total percipitation is`r round(mean(precipitation$total, na.rm = TRUE), 2)` inches on average, and the highest total monthly precipitation occured in July, 2017 with 7.09 inches.

## Problem 2_4 Answer Questions about Trash and Precipitation

*  what was the total precipitation in 2017?\
The total precipitation in 2017 are `r sum(prec_2017$total)` inches.\
</br>
* What was the median number of sports balls in a dumpster in 2016?\
The median number of sports balls in a dumpster in 2016 are `r median(filter(trash, year == 2016) %>% pull(sports_balls), na.rm = TRUE)`.

## Problem 3_1 BRFSS Data Cleaning

```{r Problem 3_1}

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("remove2", "county"), sep = 5) %>% 
  select(year, state = locationabbr, county, response, data_value) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  select(year:county, excellent, very_good, good, fair, poor) %>% 
  mutate(above_good = excellent + very_good)
  
```

## Problem 3_2 Answer Questions about BRFSS

```{r Problem 3_2}

#  What state is observed the most?
count(brfss, state) %>% 
arrange(-n) %>% 
head(1)

```

* How many unique locations are included in the dataset?\
There are `r count(distinct(brfss, state, county))` unique locations included in the dataset.\
</br>
* Is every state represented?\
Since there are `r count(distinct(brfss, state))` states and federal district in the dataset, all the states are included.\
</br>
* What state is observed the most?\
New Jersey is observed the most with 146 times.\
</br>
* In 2002, what is the median of the “Excellent” response value?\
The median of the “Excellent” response value in 2002 are `r median(filter(brfss, year == 2002) %>% pull(excellent), na.rm = TRUE)`.

## Problem 3_3 Histogram

```{r Problem 3_3}

filter(brfss, year == 2002) %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram() +
  labs(
    title = '"Excellent" response values in the year 2002',
    x = "“Excellent” response values",
    y = "Count")

```

## Problem 3_4 Scatterplot

```{r Problem 3_4}

# Make a subsets containing only New York County and Queens County
brfss_ny = filter(brfss, county == "New York County" | county == "Queens County") %>% 
  arrange(county, year)

ggplot(brfss_ny, aes(x = year, y = excellent, color = county)) + 
  geom_point(size = 4) +
  labs(
    title = "Proportion of 'Excellent' response values in New York County and Queens County each year",
    x = "Year",
    y = 'Proportion of "Excellent" response')

```






